---
title: "Analysis for Hollister, Kreakie, Milstead"
author: "Jeff Hollister, Betty Kreakie, Bryan Milstead"
date: "Monday, April 28, 2014"
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE, cache=TRUE}
devtools::install_github("hkm","jhollist")
library(hkm)
library(varSelRF)
data(hkm2014Data)  
```

## Main Question: Given that we have a unique combination of novel data, increased computational facilities, and an expanded geographic scope, what can we add to the classic understanding of the processes that are predictive of chlorophyll *a* and trophic state in lakes. 

### Initital Predictor variable selection (Jeff)
1.mean temp
2.DD45 
3.Select scale(s) - local - 300m and regional - 3000m
 *ANALYSIS*- as a check look at centroid distance matrix 
4.Turbidity, not secchi - because turb is an instrument measure and not limited by depth of lake.  If need be Turb could be converted to an estimated secchi.

We start by selecting the predictor variables. Apriori we made the following decisions:

1.  There are three water temperature measure based on the profile data for the first 2m of the water column, the mean (TmeanW), the min (TminW) and the max (TmaxW).  We decided to use TmeanW arbitrarily.

2.  There are four estimates of growing degree days:  DDs40, DDs45, DDs50, and DDs55.  The number refers to the reference temperature used in the calculation.  It really shouldn't matter for our purposes so we arbitrarily chose DD45.

3. The NLCD landcover was calculated as a percent and as total area in buffers of 300m, 1500m, and 3000m.  A fourth buffer scaled to the maximum inlake distance was also used.  We decided to use the percent landcover in the 300m (homeowner) and the 3000m buffers for the analysis.  *ANALYSIS*- as a check look at centroid distance matrix.

4. Turbidity and Secchi provide similar information and either one can be estimated from the other.  Secchi is not accurate in lakes that are clear to the bottom so we choose to use Turbidity only.

5.  We also decided to add a new variable for the NP ratio.  This is just a straight weight ratio and not the molecular weight ratio.   



```{r predictorSelection, echo=FALSE, cache=FALSE} 

	#add a variable for NP ratio
		hkm2014Data$NPratio<-hkm2014Data$NTL/hkm2014Data$PTL

	#Create a Chla Category with Low=first Quartile (Q1); Med=Q2 & Q3; High=Q4
		hkm2014Data$chlaCat<-factor('MED',ordered=T,levels=c("LOW","MED","HIGH"))
    hkm2014Data$chlaCat[hkm2014Data$CHLA<=quantile(hkm2014Data$CHLA,.25,na.rm=T)]<-'LOW'
    hkm2014Data$chlaCat[hkm2014Data$CHLA>=quantile(hkm2014Data$CHLA,.75,na.rm=T)]<-'HIGH'
    hkm2014Data$chlaCat[is.na(hkm2014Data$CHLA)]<-NA
			table(hkm2014Data$chlaCat,useNA='ifany')
		hkm2014Da$chlaCat<-chlaCat
		hkm2014Data$chlaCat<-factor(hkm2014Data$chlaCat,ordered=T,levels=c("LOW","MED","HIGH"))

	#select the predictor variable
		omit<-unique(c(grep("NLA_ID",names(hkm2014Data)),  #NLA_ID
						grep("Comid",names(hkm2014Data)),				 #Comid
						grep("NA_",names(hkm2014Data)),          #NA tabulations for Landscape variables
						grep("MaxDist",names(hkm2014Data)),      #MaxDist Buffer Variables
						grep("1500m",names(hkm2014Data)),        #1500m Buffer Variables
						grep("Km2",names(hkm2014Data)),					 #Buffer Variables tabulated as area
						grep("DDs",names(hkm2014Data))[-2],      #Degree Days variables-keep DDs45
						grep("Tm",names(hkm2014Data))[-3],       #Water Temp variables-keep TmeanW
						grep("SECMEAN",names(hkm2014Data)),  		 #SECMEAN dropped-using TURB instead
						grep("CHLA",names(hkm2014Data)),         #CHLA dropped; it is the response variable
						grep("bvCat",names(hkm2014Data)),        #bvCat: biovolume category
						grep("chlaCat",names(hkm2014Data)),      #chlaCat: new Chla category
						grep("TS_",names(hkm2014Data))))         #Trophic State Classification Variables

#create df with the predictor variables including  WSA_ECO9 as a categorical variable.	
	Pall<-hkm2014Data[,-omit]  
		#create df of All variables with 300m NLCD only
			Pall300<-Pall[,-grep("3000m",names(Pall))]
		#create df of all variables with 3000m NLCD only
			Pall3k<-Pall[,-grep("300m",names(Pall))]

#create df with the predictor variables including  WSA_ECO9 9 dummy variables.	
	#df without WSA_ECO9
		Pall1<-Pall[,-grep("WSA_ECO9",names(Pall))]
	#create dummy variables for WSA_ECO9 as factors
		a<-as.character(unique(hkm2014Data$WSA_ECO9)) #vector of WSA_ECO9 names
  		for(i in a){
    		b<-paste('WSA9_',i,sep='')   #create variable name
    		Pall1[,b]<-factor(0, levels = 0:1, labels = c("no", "yes"))  #create the factor
    		Pall1[Pall$WSA_ECO9==i,b]<-'yes'  # populate the factor
    }		

#create df of GIS predictor variables
		#variables to keep from Pall
			keep<-c(grep("_300",names(Pall),value=TRUE),"AlbersX","AlbersY","LakeArea","LakePerim","ShoreDevel",
							"BASINAREA","DEPTHMAX","ELEV_PT","DDs45","MaxLength","MaxWidth","MeanWidth",          
              "FetchN","FetchNE","FetchE","FetchSE","MaxDepthCorrect","VolumeCorrect","MeanDepthCorrect",
							grep("WSA",names(Pall),value=TRUE))
		#create df of all GIS variables
			Pgis<-Pall[,keep]
		#create df of GIS variables with 300m NLCD only
			Pgis300<-Pgis[,-grep("3000m",names(Pgis))]
		#create df of GIS variables with 3000m NLCD only
			Pgis3k<-Pgis[,-grep("300m",names(Pgis))]

```



### Analysis 1. Compare classic linear models to random forest models?

1.Chl a ~ TP (converted to TS)
2.Chl a ~ TN (converted to TS)
3.Chl a ~ TN + TP (converted to TS)
4.Chl a ~ Linear Mod with varSelRF variables (converted to TS)

use varSelRF to select the variables for linear model 4

```{r run_varSelRF, echo=FALSE, cache=TRUE} 
#run for all predictors
	#intial conditions:
		NTREE<-10000 #(10000) number of trees for first run
		Niterations<-5000 #(5000) number of trees for subsequent runs
		dropFrac<-0.2 #(0.2) proportion of variable to drop after each iteration
	#select complete cases
		keep<-complete.cases(Pall,hkm2014Data$TS_CHLA)
<<<<<<< HEAD
	#run rf.s
		print("RF of all variables with both 300m and 3000m  percent nlcd")
			varSelAll<-varSelRF(Pall[keep,],hkm2014Data[keep,'TS_CHLA'],ntree = NTREE, 
												ntreeIterat = Niterations, vars.drop.frac = dropFrac)
			#view results
				varSelAll  #Selected variables: c('NPratio,'NTL','PTL','TOC','TURB','K',WSA_ECO9)
				summary(varSelAll$selec.history)
				#plot(varSelAll, which=2)

		print("RF of all variables with only the 300m percent nlcd")
			varSelAll300<-varSelRF(Pall300[keep,],hkm2014Data[keep,'TS_CHLA'],ntree = NTREE, 
												ntreeIterat = Niterations, vars.drop.frac = dropFrac)
			#view results
				varSelAll300  #Selected variables: c(')
				summary(varSelAll300$selec.history)
		
		print("RF of all variables with only the 3000m percent nlcd")
			varSelAll3k<-varSelRF(Pall3k[keep,],hkm2014Data[keep,'TS_CHLA'],ntree = NTREE,
												ntreeIterat = Niterations, vars.drop.frac = dropFrac)
			#view results
				varSelAll3k  #Selected variables: c(')
				summary(varSelAll3k$selec.history)
=======
	#run rf
		varSelAll<-varSelRF(Pall[keep,],hkm2014Data[keep,'TS_CHLA'],ntree = 10000, 
												 ntreeIterat = 5000, vars.drop.frac = 0.2)
	#view results
		varSelAll     #Selected variables: c('NPratio,'NTL','PTL','TOC','TURB','K',WSA_ECO9)
		summary(varSelAll$selec.history)
		plot(varSelAll, which=2)
>>>>>>> 59a3329acd4b2d42233c4b46104828a02f7de5f4

#run for all GIS predictors
	#select complete cases
		keep<-complete.cases(Pgis,hkm2014Data$TS_CHLA)
	#run rf
		print("RF of GIS variables with both 300m and 3000m  percent nlcd")
			varSelGIS<-varSelRF(Pgis[keep,],hkm2014Data[keep,'TS_CHLA'],ntree = NTREE, 
												 ntreeIterat = Niterations, vars.drop.frac = dropFrac)
			#view results
				varSelGIS  #Selected variables: c("AlbersX","AlbersY","CropsPer_3000m","DEPTHMAX",
                                    #  "EvergreenPer_3000m","EvergreenPer_300m","WSA_ECO9")
				summary(varSelGIS$selec.history)
		
		print("RF of GIS variables with only 300m percent nlcd")
			varSelGIS300<-varSelRF(Pgis300[keep,],hkm2014Data[keep,'TS_CHLA'],ntree = NTREE, 
												 ntreeIterat = Niterations, vars.drop.frac = dropFrac)
			#view results
				varSelGIS300  #Selected variables: c("AlbersX","AlbersY","BASINAREA","CropsPer_300m","DDs45","DeciduousPer_300m","DEPTHMAX","ELEV_PT","EvergreenPer_300m","HerbWetPer_300m", "LakeArea","LakePerim","MaxDepthCorrect","MaxLength","MeanDepthCorrect","PasturePer_300m","VolumeCorrect","WoodyWetPer_300m","WSA_ECO9")    
				summary(varSelGIS300$selec.history)

		print("RF of GIS variables with only 3000m percent nlcd")
			varSelGIS3k<-varSelRF(Pgis3k[keep,],hkm2014Data[keep,'TS_CHLA'],ntree = NTREE, 
												 ntreeIterat = Niterations, vars.drop.frac = dropFrac)
			#view results
			varSelGIS3k  #Selected variables: c("AlbersX","AlbersY","BASINAREA","CropsPer_3000m","DDs45","DeciduousPer_3000m","DEPTHMAX","DevOpenPer_3000m","ELEV_PT","EvergreenPer_3000m","HerbWetPer_3000m","MaxDepthCorrect","MeanDepthCorrect","ShrubPer_3000m","WSA_ECO9" )
			summary(varSelGIS3k$selec.history)
```





```{r newChlaCategory, echo=FALSE, cache=FALSE}
##New varSelRF for chlaCat
#run for all predictors
	#intial conditions:
		NTREE<-10000 #(10000) number of trees for first run
		Niterations<-5000 #(5000) number of trees for subsequent runs
		dropFrac<-0.2 #(0.2) proportion of variable to drop after each iteration
	#select complete cases
		keep<-complete.cases(Pall,hkm2014Data$chlaCat)
<<<<<<< HEAD
	#run rf.s
		print("RF of all variables with both 300m and 3000m  percent nlcd")
			varSelAll<-varSelRF(Pall[keep,],hkm2014Data[keep,'chlaCat'],ntree = NTREE, 
												ntreeIterat = Niterations, vars.drop.frac = dropFrac)
			#view results
				varSelAll  #Selected variables: c('NPratio,'NTL','PTL','TOC','TURB','K',WSA_ECO9)
				summary(varSelAll$selec.history)
				#plot(varSelAll, which=2)

		print("RF of all variables with only the 300m percent nlcd")
			varSelAll300<-varSelRF(Pall300[keep,],hkm2014Data[keep,'chlaCat'],ntree = NTREE, 
												ntreeIterat = Niterations, vars.drop.frac = dropFrac)
			#view results
				varSelAll300  #Selected variables: c(')
				summary(varSelAll300$selec.history)
		
		print("RF of all variables with only the 3000m percent nlcd")
			varSelAll3k<-varSelRF(Pall3k[keep,],hkm2014Data[keep,'chlaCat'],ntree = NTREE,
												ntreeIterat = Niterations, vars.drop.frac = dropFrac)
			#view results
				varSelAll3k  #Selected variables: c(')
				summary(varSelAll3k$selec.history)
=======
	#run rf
		varSelAll<-varSelRF(Pall[keep,],hkm2014Data[keep,'chlaCat'],ntree = 10000, 
												 ntreeIterat = 5000, vars.drop.frac = 0.2)
	#view results
		varSelAll     #Selected variables: c('NPratio,'NTL','PTL','TOC','TURB','K',WSA_ECO9)
		summary(varSelAll$selec.history)
		plot(varSelAll, which=2)
>>>>>>> 59a3329acd4b2d42233c4b46104828a02f7de5f4

#run for all GIS predictors
	#select complete cases
		keep<-complete.cases(Pgis,hkm2014Data$chlaCat)
	#run rf
		print("RF of GIS variables with both 300m and 3000m  percent nlcd")
			varSelGIS<-varSelRF(Pgis[keep,],hkm2014Data[keep,'chlaCat'],ntree = NTREE, 
												 ntreeIterat = Niterations, vars.drop.frac = dropFrac)
			#view results
				varSelGIS  #Selected variables: c("AlbersX","AlbersY","CropsPer_3000m","DEPTHMAX",
                                    #  "EvergreenPer_3000m","EvergreenPer_300m","WSA_ECO9")
				summary(varSelGIS$selec.history)
		
		print("RF of GIS variables with only 300m percent nlcd")
			varSelGIS300<-varSelRF(Pgis300[keep,],hkm2014Data[keep,'chlaCat'],ntree = NTREE, 
												 ntreeIterat = Niterations, vars.drop.frac = dropFrac)
			#view results
				varSelGIS300  #Selected variables: c("AlbersX","AlbersY","BASINAREA","CropsPer_300m","DDs45","DeciduousPer_300m","DEPTHMAX","ELEV_PT","EvergreenPer_300m","HerbWetPer_300m", "LakeArea","LakePerim","MaxDepthCorrect","MaxLength","MeanDepthCorrect","PasturePer_300m","VolumeCorrect","WoodyWetPer_300m","WSA_ECO9")    
				summary(varSelGIS300$selec.history)

		print("RF of GIS variables with only 3000m percent nlcd")
			varSelGIS3k<-varSelRF(Pgis3k[keep,],hkm2014Data[keep,'chlaCat'],ntree = NTREE, 
												 ntreeIterat = Niterations, vars.drop.frac = dropFrac)
			#view results
			varSelGIS3k  #Selected variables: c("AlbersX","AlbersY","BASINAREA","CropsPer_3000m","DDs45","DeciduousPer_3000m","DEPTHMAX","DevOpenPer_3000m","ELEV_PT","EvergreenPer_3000m","HerbWetPer_3000m","MaxDepthCorrect","MeanDepthCorrect","ShrubPer_3000m","WSA_ECO9" )
			summary(varSelGIS3k$selec.history)

```


```{r linearModels, echo=FALSE, cache=FALSE} 
	#select data for linear models
		#select vars for linear model 
			LMall<-hkm2014Data[,c('CHLA','NTL','PTL','TOC','TURB','K')]
			LMall$NPratio<-Pall$NPratio
		#Log transform the data
			LMall<-log(LMall) #log transform data
				names(LMall)<-paste('ln',names(LMall),sep='')
		#add WSA9 dummies
			LMall[,7:15]<-Pall1[,grep("WSA",names(Pall1),value=TRUE)]

		
	#run linear models 
		#1. lnCHLA ~ lnNTL
				Formula<-formula(paste(names(LMall)[1],'~',paste(names(LMall)[2], collapse="+"))) 
					print(Formula)
					LM<-lm(Formula,data=LMall)
						summary(LM) 						#adjR2== 0.5231
						mean(residuals(LM)^2)		#rmse == 1.0899
		#2. lnCHLA ~ lnPTL
				Formula<- formula(paste(names(LMall)[1],'~',paste(names(LMall)[3], collapse="+"))) 
					print(Formula)
					LM<-lm(Formula,data=LMall)
						summary(LM) 						#adjR2== 0.5251
						mean(residuals(LM)^2)		#rmse == 1.0854
		#3. lnCHLA ~ lnNTL + lnPTL
				Formula<- formula(paste(names(LMall)[1],'~',paste(names(LMall)[2:3], collapse="+"))) 
					print(Formula)
					LM<-lm(Formula,data=LMall)
						summary(LM) 						#adjR2== 0.5795
						mean(residuals(LM)^2)		#rmse == 0.9600
    #4. lnCHLA ~ lnNTL + lnPTL + lnTOC + lnTURB + lnK +lnNPratio
				Formula<- formula(paste(names(LMall)[1],'~',paste(names(LMall)[2:7], collapse="+"))) 
					print(Formula)
					LM<-lm(Formula,data=LMall)
						summary(LM) 						#adjR2== 0.6603
						mean(residuals(LM)^2)		#rmse == 0.7729
		#5. lnCHLA ~ lnNTL + lnPTL + lnTOC + lnTURB + lnK + WSA9_WMT + WSA9_CPL + 
    						#WSA9_NAP + WSA9_UMW + WSA9_TPL + WSA9_SPL + WSA9_XER + WSA9_NPL + WSA9_SAP
				Formula<- formula(paste(names(LMall)[1],'~',paste(names(LMall)[-1], collapse="+"))) 
					print(Formula)
					LM<-lm(Formula,data=LMall)
						summary(LM) 						#adjR2== 0.7156
						mean(residuals(LM)^2)		#rmse == 0.6430
```


5.Chl a TS ~ RF(varSelRF variables) - single randomForest run

```{r exploreRF, echo=FALSE, cache=FALSE}
## Exploring different RF outputs
#run for all predictors
	#select complete cases
		keep<-complete.cases(Pall,hkm2014Data$TS_CHLA)

		#random forest with all our variables
		full.RF<-randomForest(hkm2014Data[keep,'TS_CHLA']~., data=Pall[keep,],ntree = 10000,
												 importance=TRUE,proximity=TRUE)

		#random forest on only varSel variables
		varSelAll.var<-c('NPratio','NTL','PTL','TOC','TURB','K','WSA_ECO9')
	
		varSelAll.RF<-randomForest(hkm2014Data[keep,'TS_CHLA']~., data=Pall[keep,varSelAll.var],ntree =10000,
												 importance=TRUE,proximity=TRUE)


		#random forest on landscape variables at 300 and 3000 scale
		keep<-complete.cases(Pgis300,hkm2014Data$TS_CHLA)
		Pgis300.RF<-randomForest(hkm2014Data[keep,'TS_CHLA']~., data=Pgis300[keep,],ntree =10000,
												 importance=TRUE,proximity=TRUE)

		keep<-complete.cases(Pgis3k,hkm2014Data$TS_CHLA)
		Pgis3k.RF<-randomForest(hkm2014Data[keep,'TS_CHLA']~., data=Pgis3k[keep,],ntree =10000,
												 importance=TRUE,proximity=TRUE)


		#random forest on varSel landscape variables at 300 and 3000 scale
		keep<-complete.cases(Pgis300,hkm2014Data$TS_CHLA)
		varSelGIS300.var<-c("AlbersX","AlbersY","BASINAREA","CropsPer_300m","DDs45","DeciduousPer_300m",
														"DEPTHMAX","ELEV_PT","EvergreenPer_300m","HerbWetPer_300m", "LakeArea","LakePerim"
														,"MaxDepthCorrect","MaxLength","MeanDepthCorrect","PasturePer_300m","VolumeCorrect"
														,"WoodyWetPer_300m","WSA_ECO9")

		Pgis300varSel.RF<-randomForest(hkm2014Data[keep,'TS_CHLA']~., data=Pgis300[keep,varSelGIS300.var],ntree =10000,
												 importance=TRUE,proximity=TRUE)

		keep<-complete.cases(Pgis3k,hkm2014Data$TS_CHLA)
		varSelGIS3k.var<-c("AlbersX","AlbersY","CropsPer_3000m","DEPTHMAX","ELEV_PT","EvergreenPer_3000m",
											 "MeanDepthCorrect","WSA_ECO9" )
		
		Pgis3kvarSel.RF<-randomForest(hkm2014Data[keep,'TS_CHLA']~., data=Pgis3k[keep,varSelGIS3k.var],ntree =10000,
												 importance=TRUE,proximity=TRUE)

		#random forest on varSel landscape variables at both scales
		keep<-complete.cases(Pgis,hkm2014Data$TS_CHLA)
	
		varSelGIS.var<- c("AlbersX","AlbersY","CropsPer_3000m","DEPTHMAX","EvergreenPer_3000m"
											,"EvergreenPer_300m","WSA_ECO9")

		PgisvarSel.RF<-randomForest(hkm2014Data[keep,'TS_CHLA']~., data=Pgis[keep,varSelGIS.var],ntree =10000,
												 importance=TRUE,proximity=TRUE)

												

```

6.Chl a TS ~ Random Forest Consensus - single randomForest run

*ANALYSIS*
- varSelRF (Bryan plus play with varSelRFBoot)
- run linear models (1-4) (Bryan)
- run RF (Betty)
 - randomForest All Variables
 - randomForest varSelRF Variables
- output error matrices (Bryan, Betty and Jeff)
 - overall accuracy
 - Kappa
 
 
```{r mob, echo=FALSE, cache=FALSE} 
#####MOB
require(party)
	#define data and transform the response variables; covariates do not need transformation.
				mobData<-data.frame(lnCHLA=log(hkm2014Data$CHLA),
												lnNTL=log(hkm2014Data$NTL),
												lnPTL=log(hkm2014Data$PTL),
												Pall1[,-c(grep("DATE_COL",names(Pall1)),
												 grep("NTL",names(Pall1)),
												 grep("PTL",names(Pall1)))])

#The model is estimated by OLS, the instability is assessed using a Bonferroni-corrected significance
#level of ? = 0.05 and the nodes are split with a required minimal segment size of 40
#observations. The control parameters are thus set to
  ctrl <- mob_control(alpha = 0.05, bonferroni = TRUE, minsplit = 20,
    objfun = deviance, verbose = F)
    
#select the covariates
  Cov<-names(mobData)[-c(1:3)]
  Formula <- formula(paste('lnCHLA~lnNTL+lnPTL |',paste(Cov, collapse="+")))  #cool way to specify the model
  #Formula <- formula(paste('lnCHLA~lnNTL |',paste(Cov, collapse="+")))  #cool way to specify the model
  
#run the model
  test <- mob(Formula,data=mobData,control=ctrl,model=linearModel)

#print results
#test

#plot results
#windows(12,9)
plot(test)

#The test statistics and p?values computed in each node, can be extracted analogously by using
#the method for the function sctest() (for performing structural change tests).
  sctest(test, node = 6)

#For summarizing the quality of the fit, we could compute the mean squared error, loglikelihood or AIC:
  mean(residuals(test)^2)  #rmse = 0.5790
  logLik(test)						 #log Lik.= -1186.26 (df=44)
  AIC(test)                #AIC = 2460.519
```


### Question 2. Are *in situ* nutrient and water quality data required to make predicitions of trophic state or can GIS derived information on lake and land use/land cover also provide quality predictions?

**Predict for these classes**
1. NLA Chl a trophic state classes
2. High/Low trophic state classes

### Question3. Can data-mining approaches identify new, interesting variables to use for the prediction of trophic state?



## Here's the analysis part:

### Initial Variable Culling:

This set of code creates namesets needed to cull out the initial set of variables

```{r variableSelection, echo=FALSE, cache=FALSE}
nlawqNames<-names(hkm2014Data)[c(155:158,160:194)]
nlawqNames<-nlawqNames[!nlawqNames%in%c("SECMEAN","TminW","TmaxW")]
gisNames<-names(hkm2014Data[c(3,149:153,195:208)])
gisNames<-gisNames[!gisNames%in%c("DD40","DD50")]
lulcPercentNames<-names(hkm2014Data[c(grep("PercentImperv",names(hkm2014Data)),grep("Per_",names(hkm2014Data)))])
if(length(grep("NA_",lulcPercentNames))>0){lulcPercentNames<-lulcPercentNames[-grep("NA_",lulcPercentNames)]}
lulcPercentNames<-lulcPercentNames[-grep("_1500m",lulcPercentNames)]
respName<-"TS_CHLA"
allVarDF<-na.omit(hkm2014Data[,c(nlawqNames,gisNames,lulcPercentNames,respName)])
```


## Chl *a* Trophic Status ~ All Variables + Landscape Percent

```{r allLandscapePercent, echo=FALSE, cache=FALSE, eval=FALSE}

allPerc_varsel<-varSelRF(allVarDF[,1:dim(allVarDF)[2]-1],allVarDF[,respName],ntree = 10000, 
												 ntreeIterat = 5000, vars.drop.frac = 0.2)
summary(allPerc_varsel$selec.history)
plot(allPerc_varsel, which=2)

```



## Chl *a* Trophic Status ~ GIS Only Variables +Landscape Percent

```{r gisLandscapePercent, echo=FALSE, cache=FALSE, eval=FALSE}
gisVarDF<-na.omit(hkm2014Data[,c(gisNames,lulcPercentNames,respName)])
gisPerc_varsel<-varSelRF(gisVarDF[,1:dim(gisVarDF)[2]-1],gisVarDF[,respName],ntree = NTREE, 
												 ntreeIterat = 5000,vars.drop.frac = 0.2)
summary(gisPerc_varsel$selec.history)
plot(gisPerc_varsel, which=2)
```



