---
title: "Analysis for Hollister, Kreakie, Milstead"
author: "Jeff Hollister, Betty Kreakie, Bryan Milstead"
date: "Monday, April 28, 2014"
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE, cache=TRUE}
devtools::install_github("hkm","jhollist")
library(hkm)
library(varSelRF)
data(hkm2014Data)  
```

## Main Question: Given that we have a unique combination of novel data, increased computational facilities, and an expanded geographic scope, what can we add to the classic understanding of the processes that are predictive of chlorophyll *a* and trophic state in lakes. 

### Initital variable selection (Jeff)
1.mean temp
2.DD45 
3.Select scale(s) - local - 300m and regional - 3000m
 *ANALYSIS*- as a check look at centroid distance matrix 
4.Turbidity, not secchi - because turb is an instrument measure and not limited by depth of lake.  If need be Turb could be converted to and estimated secchi.


```{r predictorSelection, echo=FALSE, cache=FALSE} 

	#add a variable for NP ratio
		hkm2014Data$NPratio<-hkm2014Data$NTL/hkm2014Data$PTL

	#select the predictor variable
		omit<-unique(c(grep("NLA_ID",names(hkm2014Data)),  #NLA_ID
						grep("Comid",names(hkm2014Data)),				 #Comid
						grep("NA_",names(hkm2014Data)),          #NA tabulations for Landscape variables
						grep("MaxDist",names(hkm2014Data)),      #MaxDist Buffer Variables
						grep("1500m",names(hkm2014Data)),        #1500m Buffer Variables
						grep("Km2",names(hkm2014Data)),					 #Buffer Variables tabulated as area
						grep("DDs",names(hkm2014Data))[-2],      #Degree Days variables-keep DDs45
						grep("Tm",names(hkm2014Data))[-3],       #Water Temp variables-keep TmeanW
						grep("SECMEAN",names(hkm2014Data)),  		 #SECMEAN dropped-using TURB instead
						grep("CHLA",names(hkm2014Data)),         #CHLA dropped; it is the response variable
						grep("bvCat",names(hkm2014Data)),        #bvCat: biovolume category
						grep("TS_",names(hkm2014Data))))         #Trophic State Classification Variables

#create df with the predictor variables including  WSA_ECO9 as a categorical variable.	
	Pall<-hkm2014Data[,-omit]  

#create df with the predictor variables including  WSA_ECO9 9 dummy variables.	
	#df without WSA_ECO9
		Pall1<-Pall[,-grep("WSA_ECO9",names(Pall))]
	#create dummy variables for WSA_ECO9 as factors
		a<-as.character(unique(hkm2014Data$WSA_ECO9)) #vector of WSA_ECO9 names
  		for(i in a){
    		b<-paste('WSA9_',i,sep='')   #create variable name
    		Pall1[,b]<-factor(0, levels = 0:1, labels = c("no", "yes"))  #create the factor
    		Pall1[Pall$WSA_ECO9==i,b]<-'yes'  # populate the factor
    }		

#create df of GIS predictor variables
		#variables to keep from Pall
			keep<-c(grep("_300",names(Pall),value=TRUE),"AlbersX","AlbersY","LakeArea","LakePerim","ShoreDevel",
							"BASINAREA","DEPTHMAX","ELEV_PT","DDs45","MaxLength","MaxWidth","MeanWidth",          
              "FetchN","FetchNE","FetchE","FetchSE","MaxDepthCorrect","VolumeCorrect","MeanDepthCorrect",
							grep("WSA",names(Pall),value=TRUE))
		#create df of all GIS variables
			Pgis<-Pall[,keep]
		#create df of GIS variables with 300m NLCD only
			Pgis300<-Pgis[,-grep("3000m",names(Pgis))]
		#create df of GIS variables with 3000m NLCD only
			Pgis3k<-Pgis[,-grep("300m",names(Pgis))]

```



### Analysis 1. Compare classic linear models to random forest models?

1.Chl a ~ TP (converted to TS)
2.Chl a ~ TN (converted to TS)
3.Chl a ~ TN + TP (converted to TS)
4.Chl a ~ Linear Mod with varSelRF variables (converted to TS)

use varSelRF to select the variables for linear model 4

```{r run_varSelRF, echo=FALSE, cache=TRUE} 
#run for all predictors
	#select complete cases
		keep<-complete.cases(Pall,hkm2014Data$TS_CHLA)
	#run rf
		varSelAll<-varSelRF(Pall[keep,],hkm2014Data[keep,'TS_CHLA'],ntree = 10000, 
												 ntreeIterat = 5000, vars.drop.frac = 0.2)
	#view results
		varSelAll  #Selected variables: c('NPratio,'NTL','PTL','TOC','TURB','K',WSA_ECO9)
		summary(varSelAll$selec.history)
		plot(varSelAll, which=2)

#run for all GIS predictors
	#select complete cases
		keep<-complete.cases(Pgis,hkm2014Data$TS_CHLA)
	#run rf
		varSelGIS<-varSelRF(Pgis[keep,],hkm2014Data[keep,'TS_CHLA'],ntree = 10000, 
												 ntreeIterat = 5000, vars.drop.frac = 0.2)
	#view results
		varSelGIS  #Selected variables: c("AlbersX","AlbersY","CropsPer_3000m","DEPTHMAX",
                                    #  "EvergreenPer_3000m","EvergreenPer_300m","WSA_ECO9")
		summary(varSelGIS$selec.history)
		plot(varSelGIS, which=2)

#run for GIS predictors with the 300m NLCD but without the 3000m NLCD
	#select complete cases
		keep<-complete.cases(Pgis300,hkm2014Data$TS_CHLA)
	#run rf
		varSelGIS300<-varSelRF(Pgis300[keep,],hkm2014Data[keep,'TS_CHLA'],ntree = 10000, 
												 ntreeIterat = 5000, vars.drop.frac = 0.2)
	#view results
		varSelGIS300  #Selected variables: c("AlbersX","AlbersY","BASINAREA","CropsPer_300m","DDs45","DeciduousPer_300m","DEPTHMAX","ELEV_PT","EvergreenPer_300m","HerbWetPer_300m", "LakeArea","LakePerim","MaxDepthCorrect","MaxLength","MeanDepthCorrect","PasturePer_300m","VolumeCorrect","WoodyWetPer_300m","WSA_ECO9"    
		summary(varSelGIS300$selec.history)
		plot(varSelGIS300, which=2)

#run for GIS predictors with the 3000m NLCD but without the 300m NLCD
	#select complete cases
		keep<-complete.cases(Pgis3k,hkm2014Data$TS_CHLA)
	#run rf
		varSelGIS3k<-varSelRF(Pgis3k[keep,],hkm2014Data[keep,'TS_CHLA'],ntree = 10000, 
												 ntreeIterat = 5000, vars.drop.frac = 0.2)
	#view results
		varSelGIS3k  #Selected variables: c("AlbersX","AlbersY","BASINAREA","CropsPer_3000m","DDs45","DeciduousPer_3000m","DEPTHMAX","DevOpenPer_3000m","ELEV_PT","EvergreenPer_3000m","HerbWetPer_3000m","MaxDepthCorrect","MeanDepthCorrect","ShrubPer_3000m","WSA_ECO9" )
		summary(varSelGIS3k$selec.history)
		plot(varSelGIS3k, which=2)
```


```{r linearModels, echo=FALSE, cache=FALSE} 
	#select data for linear models
		#select vars for linear model 
			LMall<-hkm2014Data[,c('CHLA','NTL','PTL','TOC','TURB','K')]
			LMall$NPratio<-Pall$NPratio
		#Log transform the data
			LMall<-log(LMall) #log transform data
				names(LMall)<-paste('ln',Vars,sep='')
		#add WSA9 dummies
			LMall[,7:15]<-Pall1[,grep("WSA",names(Pall1),value=TRUE)]

		
	#run linear models 
		#1. lnCHLA ~ lnNTL
				Formula<-formula(paste(names(LMall)[1],'~',paste(names(LMall)[2], collapse="+"))) 
					LM<-lm(Formula,data=LMall)
						summary(LM) 						#adjR2== 0.5231
						mean(residuals(LM)^2)		#rmse == 1.0899
		#2. lnCHLA ~ lnPTL
				Formula<- formula(paste(names(LMall)[1],'~',paste(names(LMall)[3], collapse="+"))) 
					LM<-lm(Formula,data=LMall)
						summary(LM) 						#adjR2== 0.5251
						mean(residuals(LM)^2)		#rmse == 1.0854
		#3. lnCHLA ~ lnNTL + lnPTL
				Formula<- formula(paste(names(LMall)[1],'~',paste(names(LMall)[2:3], collapse="+"))) 
					LM<-lm(Formula,data=LMall)
						summary(LM) 						#adjR2== 0.5795
						mean(residuals(LM)^2)		#rmse == 0.9600
    #4. lnCHLA ~ lnNTL + lnPTL + lnTOC + lnTURB + lnK +lnNPratio
				Formula<- formula(paste(names(LMall)[1],'~',paste(names(LMall)[2:7], collapse="+"))) 
					LM<-lm(Formula,data=LMall)
						summary(LM) 						#adjR2== 0.6603
						mean(residuals(LM)^2)		#rmse == 0.7729
		#5. lnCHLA ~ lnNTL + lnPTL + lnTOC + lnTURB + lnK + WSA9_WMT + WSA9_CPL + 
    						#WSA9_NAP + WSA9_UMW + WSA9_TPL + WSA9_SPL + WSA9_XER + WSA9_NPL + WSA9_SAP
				Formula<- formula(paste(names(LMall)[1],'~',paste(names(LMall)[-1], collapse="+"))) 
					LM<-lm(Formula,data=LMall)
						summary(LM) 						#adjR2== 0.7156
						mean(residuals(LM)^2)		#rmse == 0.6430
```


5.Chl a TS ~ RF(varSelRF variables) - single randomForest run
6.Chl a TS ~ Random Forest Consensus - single randomForest run

*ANALYSIS*
- varSelRF (Bryan plus play with varSelRFBoot)
- run linear models (1-4) (Bryan)
- run RF (Betty)
 - randomForest All Variables
 - randomForest varSelRF Variables
- output error matrices (Bryan, Betty and Jeff)
 - overall accuracy
 - Kappa
 
 
```{r mob, echo=FALSE, cache=FALSE} 
#####MOB
require(party)
	#define data and transform the response variables; covariates do not need transformation.
				mobData<-data.frame(lnCHLA=log(hkm2014Data$CHLA),
												lnNTL=log(hkm2014Data$NTL),
												lnPTL=log(hkm2014Data$PTL),
												Pall1[,-c(grep("DATE_COL",names(Pall1)),
												 grep("NTL",names(Pall1)),
												 grep("PTL",names(Pall1)))])

#The model is estimated by OLS, the instability is assessed using a Bonferroni-corrected significance
#level of ? = 0.05 and the nodes are split with a required minimal segment size of 40
#observations. The control parameters are thus set to
  ctrl <- mob_control(alpha = 0.05, bonferroni = TRUE, minsplit = 20,
    objfun = deviance, verbose = TRUE)
    
#select the covariates
  Cov<-names(mobData)[-c(1:3)]
  Formula <- formula(paste('lnCHLA~lnNTL+lnPTL |',paste(Cov, collapse="+")))  #cool way to specify the model
  #Formula <- formula(paste('lnCHLA~lnNTL |',paste(Cov, collapse="+")))  #cool way to specify the model
  
#run the model
  test <- mob(Formula,data=mobData,control=ctrl,model=linearModel)

#print results
#test

#plot results
#windows(12,9)
plot(test)

#The test statistics and p?values computed in each node, can be extracted analogously by using
#the method for the function sctest() (for performing structural change tests).
  sctest(test, node = 6)

#For summarizing the quality of the fit, we could compute the mean squared error, loglikelihood or AIC:
  mean(residuals(test)^2)  #rmse = 0.5790
  logLik(test)						 #log Lik.= -1186.26 (df=44)
  AIC(test)                #AIC = 2460.519
```


### Question 2. Are *in situ* nutrient and water quality data required to make predicitions of trophic state or can GIS derived information on lake and land use/land cover also provide quality predictions?

**Predict for these classes**
1. NLA Chl a trophic state classes
2. High/Low trophic state classes

### Question3. Can data-mining approaches identify new, interesting variables to use for the prediction of trophic state?



## Here's the analysis part:

### Initial Variable Culling:

This set of code creates namesets needed to cull out the initial set of variables

```{r variableSelection, echo=FALSE, cache=FALSE}
nlawqNames<-names(hkm2014Data)[c(155:158,160:194)]
nlawqNames<-nlawqNames[!nlawqNames%in%c("SECMEAN","TminW","TmaxW")]
gisNames<-names(hkm2014Data[c(3,149:153,195:208)])
gisNames<-gisNames[!gisNames%in%c("DD40","DD50")]
lulcPercentNames<-names(hkm2014Data[c(grep("PercentImperv",names(hkm2014Data)),grep("Per_",names(hkm2014Data)))])
if(length(grep("NA_",lulcPercentNames))>0){lulcPercentNames<-lulcPercentNames[-grep("NA_",lulcPercentNames)]}
lulcPercentNames<-lulcPercentNames[-grep("_1500m",lulcPercentNames)]
respName<-"TS_CHLA"
allVarDF<-na.omit(hkm2014Data[,c(nlawqNames,gisNames,lulcPercentNames,respName)])
```


## Chl *a* Trophic Status ~ All Variables + Landscape Percent

```{r allLandscapePercent, echo=FALSE, cache=FALSE, eval=FALSE}

allPerc_varsel<-varSelRF(allVarDF[,1:dim(allVarDF)[2]-1],allVarDF[,respName],ntree = 10000, 
												 ntreeIterat = 5000, vars.drop.frac = 0.2)
summary(allPerc_varsel$selec.history)
plot(allPerc_varsel, which=2)

```



## Chl *a* Trophic Status ~ GIS Only Variables +Landscape Percent

```{r gisLandscapePercent, echo=FALSE, cache=FALSE, eval=FALSE}
gisVarDF<-na.omit(hkm2014Data[,c(gisNames,lulcPercentNames,respName)])
gisPerc_varsel<-varSelRF(gisVarDF[,1:dim(gisVarDF)[2]-1],gisVarDF[,respName],ntree = 10000, 
												 ntreeIterat = 5000,vars.drop.frac = 0.2)
summary(gisPerc_varsel$selec.history)
plot(gisPerc_varsel, which=2)
```



