---
title: "Analysis for Hollister, Kreakie, Milstead"
author: "Jeff Hollister, Betty Kreakie, Bryan Milstead"
date: "Monday, April 28, 2014"
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE, cache=TRUE}
devtools::install_github("hkm","jhollist")
library(hkm)
library(varSelRF)
data(hkm2014Data)  
```

## Main Question: Given that we have a unique combination of novel data, increased computational facilities, and an expanded geographic scope, what can we add to the classic understanding of the processes that are predictive of chlorophyll *a* and trophic state in lakes. 

### Initital Predictor variable selection 
We start by selecting the predictor variables. *A priori* we made the following decisions:

1.  There are three water temperature measure based on the profile data for the first 2m of the water column, the mean (TmeanW), the min (TminW) and the max (TmaxW).  We decided to use TmeanW arbitrarily.

2.  There are four estimates of growing degree days:  DDs40, DDs45, DDs50, and DDs55.  The number refers to the reference temperature used in the calculation.  It really shouldn't matter for our purposes so we arbitrarily chose DD45.

3. The NLCD landcover was calculated as a percent and as total area in buffers of 300m, 1500m, and 3000m.  A fourth buffer scaled to the maximum inlake distance was also used.  We decided to use the percent landcover in the 300m (homeowner) and the 3000m buffers for the analysis.  *ANALYSIS*- as a check look at centroid distance matrix.

4. Turbidity and Secchi provide similar information and either one can be estimated from the other.  Secchi is not accurate in lakes that are clear to the bottom so we choose to use Turbidity only.

5.  We also decided to add a new variable for the NP ratio.  This is just a straight weight ratio and not the molecular weight ratio.  

6.  If the WSA_ECO9 is a categorical variable with the names of the nine "Wadeable Stream Assessment" aggregated ecoregions.  If this variable is selected it will need to be converted to dummy variable for the linear analysis.  A new data.frame call "WSA9dummies" is created for this.

This give us one data.frame:

1.  "Pall" contains the predictor variables (including the categorical variable "WSA_ECO9")

```{r predictorSelection, echo=FALSE, cache=FALSE} 
#create the data.frame of predictors
	#It is easiest to select variables in HKM2014Data that we don't want and then delete them.  
		omit<-unique(c(grep("NLA_ID",names(hkm2014Data)),  #NLA_ID
						grep("Comid",names(hkm2014Data)),				 #Comid
						grep("NA_",names(hkm2014Data)),          #NA tabulations for Landscape variables
						grep("MaxDist",names(hkm2014Data)),      #MaxDist Buffer Variables
						grep("1500m",names(hkm2014Data)),        #1500m Buffer Variables
						grep("Km2",names(hkm2014Data)),					 #Buffer Variables tabulated as area
						grep("DDs",names(hkm2014Data))[-2],      #Degree Days variables-keep DDs45
						grep("Tm",names(hkm2014Data))[-3],       #Water Temp variables-keep TmeanW
						grep("SECMEAN",names(hkm2014Data)),  		 #SECMEAN dropped-using TURB instead
						grep("CHLA",names(hkm2014Data)),         #CHLA dropped; it is the response variable
						grep("bvCat",names(hkm2014Data)),        #bvCat: biovolume category
						grep("TS_",names(hkm2014Data))))         #Trophic State Classification Variables

		#create df by omitting the unwanted variables	
			Pall<-hkm2014Data[,-omit]  

		#Add a variable for NP ratio
			Pall$NPratio<-Pall$NTL/Pall$PTL
```

### Initital GIS Predictor variable selection 
The Predictor variables in Pall contains a mixture of variables that can be assessed with GIS and  measured water quality variables.  To see if the GIS variables alone can be used we create a new data.frame with just the GIS predictors.

1.  This creates one df named "Pgis" that includes the location, lake morphmetrics, growing degreee days (DDs45), and the NLCD percents in buffers of 300m and 3000m.
2.  You can subset this to use buffers of 300m only with this code: Pgis300<-Pgis[,-grep("3000m",names(Pgis))]   
3.  You can subset this to use buffers of 3000m only with this code: Pgis3k<-Pgis[,-grep("300m",names(Pgis))]


```{r GISpredictors, echo=FALSE, cache=FALSE} 
#create the data.frame of GIS predictors
	#Select variables from Pall to include  
			keep<-c(grep("_300",names(Pall),value=TRUE),"AlbersX","AlbersY","LakeArea","LakePerim","ShoreDevel",
							"BASINAREA","DEPTHMAX","ELEV_PT","DDs45","MaxLength","MaxWidth","MeanWidth",          
              "FetchN","FetchNE","FetchE","FetchSE","MaxDepthCorrect","VolumeCorrect","MeanDepthCorrect",
							grep("WSA",names(Pall),value=TRUE))
		#create df of all GIS variables
			Pgis<-Pall[,keep]
```

### response variable selection 
For the Random Forest analysis we need a response variable.  For now just just the TS_CHLA. (5 missing values) 


```{r responseSelection, echo=FALSE, cache=FALSE} 
	TS_CHLA<-hkm2014Data$TS_CHLA
	table(TS_CHLA,useNA='ifany') #note there are five missing values
```

### Analysis1:  Run varSelRF multiple time to see if the same variables are selected each time.


```{r run_varSelRF, echo=FALSE, cache=TRUE} 
#run for all GIS predictors
	#intial conditions:
		NTREE<-1000 #(10000) number of trees for first run
		Niterations<-500 #(5000) number of trees for subsequent runs
		dropNum<-1 # number of variables to drop after each iteration
		numRun<-10 #number of times to repeat the rf
	
	#select complete cases
		keep<-complete.cases(Pgis,TS_CHLA)

	#run rf
		Out<-data.frame(Model=rep('xxx',numRun))
			for(i in c(1:numRun)){
				rfGIS<-varSelRF(Pgis[keep,],TS_CHLA[keep],ntree=NTREE,ntreeIterat=Niterations,
												vars.drop.frac=NULL,vars.drop.num=dropNum)
		Out[i,]<-rfGIS$selected.model
		}

```

### Analysis2:  Alternatively we could use varSelRFBoot to bootstrap the analysis


```{r run_varSelRFBoot, echo=FALSE, cache=TRUE} 
#run for all GIS predictors
	#intial conditions:
		NTREE<-100 #(1000) number of trees for first run
		Niterations<-50 #(500) number of trees for subsequent runs
		dropNum<-2 # number of variables to drop after each iteration
		numBoot<-2 #number of times to bootrstrap the rf
	
	#select complete cases
		#keep<-complete.cases(Pgis,TS_CHLA)
		keep<-c(1:100)
	
	#run rfBoot
		rfGIS<-varSelRF(Pgis[keep,],TS_CHLA[keep],ntree=NTREE,ntreeIterat=Niterations,
												vars.drop.frac=NULL,vars.drop.num=dropNum)  #first run sets initial conditions
		bootGIS<-varSelRFBoot(Pgis[keep,],TS_CHLA[keep],vars.drop.num=dropNum,
											 bootnumber=numBoot,usingCluster = FALSE,srf=rfGIS)

		bootGIS
		summary(bootGIS)
		plot(bootGIS)
```


