---
title: "Working version of Random Forests For Paper"
author: "Jeff Hollister"
date: "Wednesday, April 23, 2014"
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE, cache=FALSE}
library(hkm)
library(varSelRF)
data(hkm2014Data)
```

This document represents the various modeling efforts we have tried for the HKM2014 paper.  First step in this analyis is to set up the groups of variables.  Four groups, plus the Trophic State response variable were initially indentified.  

First, are the water quality variables from NLA.

```{r nlaWaterQual,echo=FALSE, cache=FALSE}
nlawqNames<-names(hkm2014Data)[c(155:158,160:194)]
nlawqNames
```

Second,are the GIS based variables

```{r gisVariables, echo=FALSE, cache=FALSE}
gisNames<-names(hkm2014Data[c(3,149:153,195:208)])
gisNames
```

Third, are the landscape variables based on total area

```{r lulcVariablesArea, echo=FALSE, cache=FALSE}
lulcAreaNames<-names(hkm2014Data[grep("Km2_",names(hkm2014Data))])
if(length(grep("NA_",lulcAreaNames))>0){lulcAreaNames<-lulcAreaNames[-grep("NA_",lulcAreaNames)]}
lulcAreaNames
```

Fourth, the percent landscape variables (also inlcluding percent impervious) 
```{r lulcVariablesPerc, echo=FALSE, cache=FALSE}
lulcPercentNames<-names(hkm2014Data[c(grep("PercentImperv",names(hkm2014Data)),grep("Per_",names(hkm2014Data)))])
if(length(grep("NA_",lulcPercentNames))>0){lulcPercentNames<-lulcPercentNames[-grep("NA_",lulcPercentNames)]}
lulcPercentNames
```

Lastly, the response variable

```{r response, echo=FALSE, cache=FALSE}
respName<-"TS_CHLA"
respName
```

## Chl *a* Trophic Status ~ All Variables + Landscape Total Area

This first random forest predicts Cholorophyl Trophic status using all NLA variables, all GIS variables except for the landscape variables summarized by total area.

```{r allLandscapeArea, echo=FALSE, cache=FALSE}
allVarDF<-na.omit(hkm2014Data[,c(nlawqNames,gisNames,lulcAreaNames,respName)])
allArea_varsel<-varSelRF(allVarDF[,1:dim(allVarDF)[2]-1],allVarDF[,respName],ntree = 10000, ntreeIterat = 5000,vars.drop.frac = 0.2)
summary(allArea_varsel$selec.history)
plot(allArea_varsel, which=2)
```


## Chl *a* Trophic Status ~ All Variables + Landscape Percent

Our second model uses all NLA and GIS variables, and landscape var summarized by total percent.

```{r allLandscapePercent, echo=FALSE, cache=FALSE}
allVarDF<-na.omit(hkm2014Data[,c(nlawqNames,gisNames,lulcPercentNames,respName)])
allPerc_varsel<-varSelRF(allVarDF[,1:dim(allVarDF)[2]-1],allVarDF[,respName],ntree = 10000, ntreeIterat = 5000,vars.drop.frac = 0.2)
summary(allPerc_varsel$selec.history)
plot(allPerc_varsel, which=2)
```


## Chl *a* Trophic Status ~ GIS Only Variables + Landscape Total Area

This third model uses only GIS derived variables with landscape summarized by total area

```{r gisLandscapeArea, echo=FALSE, cache=FALSE}
gisVarDF<-na.omit(hkm2014Data[,c(gisNames,lulcAreaNames,respName)])
gisArea_varsel<-varSelRF(gisVarDF[,1:dim(gisVarDF)[2]-1],gisVarDF[,respName],ntree = 10000, ntreeIterat = 5000,vars.drop.frac = 0.2)
summary(gisArea_varsel$selec.history)
plot(gisArea_varsel, which=2)
```


## Chl *a* Trophic Status ~ GIS Only Variables +Landscape Percent

This third model uses only GIS derived variables with landscape summarized by total area

```{r gisLandscapePercent, echo=FALSE, cache=FALSE}
gisVarDF<-na.omit(hkm2014Data[,c(gisNames,lulcPercentNames,respName)])
gisPerc_varsel<-varSelRF(gisVarDF[,1:dim(gisVarDF)[2]-1],gisVarDF[,respName],ntree = 10000, ntreeIterat = 5000,vars.drop.frac = 0.2)
summary(gisPerc_varsel$selec.history)
plot(gisPerc_varsel, which=2)
```
